---
title: "Exercise 3"
subtitle: "Fortgeschrittene Statistische Software für NF - SS 2022/23"
author: "Nele Steiner"
date: "`r Sys.Date()`"
---

## Exercise 1: Initializing git (4 Points)

For this whole exercise sheet we will be tracking all our changes to it
in git.

a)  Start by initializing a new R project with git support, called
    `exeRcise-sheet-3`. If you forgot how to do this, you can follow
    this
    [guide](https://malikaihle.github.io/Introduction-RStudio-Git-GitHub/rstudio_project.html).
b)  Commit the files generated by Rstudio.
c)  For all of the following tasks in this exercise sheet we ask you to
    always commit your changes after finishing each subtask e.g. create
    a commit after task *1d*, *1e* etc.

> Note: This applies only to answers that have text or code as their
> answer. If you complete tasks in a different order or forget to commit
> one, this is no problem. If you change your answers you can just
> create multiple commits to track the changes.

d)  Name 2 strengths and 2 weaknesses of git. (Don't forget to create a
    commit after this answer, see *1c*)
    
Strengths:
1. Git looks at the changes in your files and records them over time to give you a full history. 
2. Git allows easy collaboration as it allows everyone who is working on a project to see the changes other persons did. 

Weaknesses:
1. Git is, especially at first complicated to use. 
2. Git needs to be used explicitly. It doesn`t just work in the background. 

e)  Knit this exercise sheet. Some new files will automatically be
    generated when knitting the sheet e.g. the HTML page. Ignore these
    files, as we only want to track the source files themselves.
    
The only file which was knitted is a HTML page. 

## Exercise 2: Putting your Repository on GitHub (3.5 Points)

For this task you will upload your solution to GitHub.

a)  Create a new repository on GitHub in your account named
    `exeRcise-sheet-3`. Make sure you create a **public repository** so
    we are able to see it for grading. Add the link to the repository
    below:
    git@github.com:Nele-md/exeRcise-sheet-3.git
b)  Push your code to this new repository by copying and executing the
    snippet on github listed under
    `…or push an existing repository from the command line`.
c)  Regularly push your latest changes to GitHub again and especially do
    so when you are finished with this sheet.

## Exercise 3: Baby-Names in Munich (4.5 Points)

Download the latest open datasets on given names ("Vornamen") from the
open data repository of the city of Munich for the years `2022` and
`2021`.

Link: <https://opendata.muenchen.de/dataset/vornamen-von-neugeborenen>

a)  Download the data for both years and track it in git. For small
    datasets like these adding them to git is not a problem.

b)  Load the data for both years into R. Check the type of the count
    variable ("Anzahl") and look into the data to determine why it is
    not numeric? Fix the problem in an appropriate manner, it is OK if
    some of the counts are inaccurate because of this. Explain your
    solution and the repercussions.
```{r}
vornamen_2021 <- read.csv("data/vornamen_2021.csv")
vornamen_2022 <- read.csv("data/open_data_portal_2022.csv")

vornamen_2021 <- vornamen_2021 %>% 
  rename(first_name = Vorname,
    number = Anzahl,
    gender = Geschlecht)

vornamen_2022 <- vornamen_2022 %>% 
  rename(first_name = Vorname,
    number = Anzahl,
    gender = Geschlecht)

typeof(vornamen_2021$Anzahl)
typeof(vornamen_2022$Anzahl)

vornamen_2021$Anzahl <- as.numeric(vornamen_2021$Anzahl)
vornamen_2022$Anzahl <- as.numeric(vornamen_2022$Anzahl)

vornamen_2021_without_NAs <- na.omit(vornamen_2021)
vornamen_2022_without_NAs <- na.omit(vornamen_2022)
```
The variable "Anzahl" is a character. It is not numeric as their are multiple NAs in the dataset. 
"as.numeric" changes the datatyp from character to numeric or rather to double, which is a typ of numeric data. 
Note: NAs introduced by coersion 
The missing values are deleted because otherwise it is not possible to calculate with the variable "Anzahl", as required by task 3c), for example. 

c)  Calculate the total number of babies born in Munich in 2022
    and 2021. Which year had the bigger baby-boom?
```{r}
total_number_of_babies_2021 <- sum(vornamen_2021_without_NAs$Anzahl)
total_number_of_babies_2022 <- sum(vornamen_2022_without_NAs$Anzahl)

total_number_of_babies_2021
total_number_of_babies_2022
```
The year 2021 had the bigger baby-boom with 11524 babies in total compared to the 9899 babies that where born in the year after. 

d)  Add a new column `year` to both datasets which holds the correct
    year for each.
```{r}
vornamen_2021_without_NAs$year <- 2021
vornamen_2022_without_NAs$year <- 2022
```


e)  Combine both datasets into one using `bind_rows()`.
```{r}
library(dplyr)
combination_2021_2022 <- bind_rows(vornamen_2021_without_NAs, vornamen_2022_without_NAs)
```

f)  Combine the counts for same names to determine the most popular
    names across both years. Print out the top 10 names in a nicely
    formatted table for both years. Include a table caption.
```{r}
most_popular_names <- 
  %>% 
  arrange(desc(Anzahl))

head(most_popular_names, 10)
```


## Exercise 4: Chat GPT + apply (3 points)

For this task: Specifically use ChatGPT to solve the task and submit
your prompts in addition to the solution

a)  The code below does not work because the wrong apply function has
    been used. Find out which apply function would be correct and why it
    did not work. Correct the code. Also calculate the rowwise means.

###Create a sample data frame

tax_data \<- data.frame( 
Name = c("Munich GmbH", "ABC Inc.", "Backpacks 1980", "Bavarian Circus"), 
Tax_2019 = c(5000, 4000, 6000, 3500),
Tax_2020 = c(4800, 4200, 5800, 3700), 
Tax_2021 = c(5200, 3800, 5900, 3400) 
)

###Calculate column-wise means

column_means \<- lapply(tax_data\[, -1\], 2, mean)

column_means

```{r}
# Corrected code by chat gbt
# Create a sample data frame
tax_data <- data.frame(
  Name = c("Munich GmbH", "ABC Inc.", "Backpacks 1980", "Bavarian Circus"),
  Tax_2019 = c(5000, 4000, 6000, 3500),
  Tax_2020 = c(4800, 4200, 5800, 3700),
  Tax_2021 = c(5200, 3800, 5900, 3400)
)

# Calculate column-wise means
column_means <- lapply(tax_data[, -1], mean)

column_means

```
There where two issues in the original code: 
1. The assignment operator "<-" was incorrectly written as "\<-". The correct operator is <-, which is used for assigning values to variables in R.

2. The "lapply" function call had an incorrect argument. Instead of 2, you need to pass the mean function as the second argument. So the correct call should be "lapply(tax_data[, -1], mean)".


b)  Using ChatGPT try to understand what the rapply() function does.
    Create an easy example with mock data where the function is used and
    explain it in your words.
Explanation: The "rapply" function applies a function to each element of a list or nested list structure, and it traverses the structure recursively, applying the function to all levels of nesting.
```{r}
# Easy example 
my_list <- list(a = 1:3, b = 4:6, c = 7:9)
rapply(my_list, function(x) x^2)
```
My explanation: In this example, "rapply" is used to take each value of "my_list" to the power of two.

## Final Note

Make sure to push all your commits and changes to GitHub before
submittining the exercise sheet.
